---
title: "Tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Installation

```{r setup, eval = FALSE}
### Optionally: specify GitHub Personal Access Token (GITHUB_PAT)
### See here why this might be important for you:
### https://kwb-r.github.io/kwb.pkgbuild/articles/install.html#set-your-github_pat

# Sys.setenv(GITHUB_PAT = "mysecret_access_token")

# Install package "remotes" from CRAN
if (! require("remotes")) {
  install.packages("remotes", repos = "https://cloud.r-project.org")
}

# Install KWB package 'kwb.heatsine' from GitHub
remotes::install_github("KWB-R/kwb.heatsine")
```

# Data prerequisites 

The sinus fit optimisation works for for daily time series of a surface water
and a groundwater monitoring point, which follow the following requirements:

## File naming

The time series must be provided in a `.csv` (comma separated value) file, which
follows the naming convention:

temperature_`<type>`_`<monitoring_id>`.csv

where `<type>` needs to be exactly `groundwater` or `surface-water`. In case any
other value for `<type>` is chosen the program will not work (e.g `SURFACE
WATER`). The value of the `<monitoring_id>` will be used for labeling the
monitoring points in the plots and can be almost any value but must not contain
`_` (underscore) or `" "` (space).

## Data structure  

The data structure of each `.csv` file is very simple as it contains only the
two columns `date` (formatted: `YYYY-MM-DD`, e.g. `2020-06-23`) and `value`
(temperature data in degree Celsius) as shown exemplary below:

```{r eval = FALSE}
date,value
2020-06-23,12.3
2020-06-24,12.3
```

Please make sure that both columns `date` and `value` are all lowercase and
correctly spelled as the program will not work if this is not the case.

# Data preparation

## Import

```{r}
## Load the R package
library(kwb.heatsine)

load_temp <- function(base_name) {
  kwb.heatsine::load_temperature_from_csv(
    kwb.heatsine::extdata_file(base_name)
  )
}

data_sw <- load_temp("temperature_surface-water_TEGsee-mikrosieb.csv")
data_gw <- load_temp("temperature_groundwater_TEG343.csv")
```

## Visualise

Plot time series for whole time period (moveover on data points to select)
  
### Surface water

```{r}
kwb.heatsine::plot_temperature_interactive(data_sw)
```

### Groundwater

```{r}
kwb.heatsine::plot_temperature_interactive(data_gw)
```

## Select

Reduce time period to a full sinus period which is definde by the user input 
(`date_end` is optional, if not given it is set to 'date_start' + 365.25 days).

```{r}
data_sw_selected <- kwb.heatsine::select_timeperiod(
  data_sw, 
  date_start = "2015-10-10",
  date_end = "2016-10-14"
)

data_gw_selected <- kwb.heatsine::select_timeperiod(
  data_gw,
  date_start = "2015-12-28",
  date_end = "2016-12-26"
)
```

Plot selected datasets:

### Surface Water (selected)

```{r}
kwb.heatsine::plot_temperature_interactive(df = data_sw_selected)
```

### Grundwater (selected)

```{r}
kwb.heatsine::plot_temperature_interactive(df = data_gw_selected)
```

# Sinus optimisation 


## Run

Sinus fit operation is performed by minimizing the RMSE error between observed 
and simulated temperature curves.

```{r}
limits <- c(100, 500) # minimum/maximum period length
tolerance <- 0.001 # the desired accuracy ()
debug <- TRUE

# Helper function to do the sinus optimisation
do_sinus_optimisation <- function(temp_df) {
  kwb.heatsine::optimise_sinus_variablePeriod(
    temp_df = temp_df,
    opt_limits = limits,
    opt_tolerance = tolerance,
    opt_debug = debug
  )
}

sinusfit_sw <- do_sinus_optimisation(temp_df = data_sw_selected)
sinusfit_gw <- do_sinus_optimisation(temp_df = data_gw_selected)

# Generate data frame with simulated and observed values
predictions <- kwb.heatsine::get_predictions(
  sinusfit_sw = sinusfit_sw, 
  sinusfit_gw = sinusfit_gw, 
  retardation_factor = 1.8
)
```

## Analyse

### Interactive Plot

```{r}
kwb.heatsine::plot_prediction_interactive(predictions)
```

### Table: Traveltimes
```{r}
knitr::kable(predictions$traveltimes)
```

### Table: Goodness of Fit 

While only RMSE is used for optimising the sinus fit, a number of goodness-of-fit 
parameters is also calculated and shown in the table below. For details on each 
of these parameters the use is referred to the function `hydroGOF::gof()` which is 
used for calculating all these values.

```{r}
knitr::kable(predictions$gof)
```

### Table: Optimisation Parameters

```{r}
knitr::kable(predictions$paras)
```

### Table: Data

```{r}
# only first entries
knitr::kable(head(predictions$data))
```

## Export 

Export results to `csv`:

```{r eval = FALSE}
write_to_csv <- function(df, file_base) {
  readr::write_csv(df, path = file.path(
    kwb.utils::createDirectory("csv"), paste0(file_base, ".csv")
  ))
}

get_residuals <- function(df) {
  data.frame(residuals = df$simulated - df$observed)
}

write_to_csv(predictions$data, "sinus-fit_predictions")
write_to_csv(predictions$paras, "sinus-fit_parameters")
write_to_csv(predictions$gof, "sinus-fit_goodness-of-fit")
write_to_csv(predictions$traveltimes, "sinus-fit_traveltimes")
write_to_csv(get_residuals(predictions$data), "sinus-fit_residuals")

## Export plots:

save_to_html <- function(p, file_base) {
  htmlwidgets::saveWidget(p, paste0(file_base, ".html"))
}

withr::with_dir(new = kwb.utils::createDirectory("plots"), code = {
  
  save_to_html(
    plot_temperature_interactive(data_sw), 
    "temperature_surface-water_time-series_full"
  )
  
  save_to_html(
    plot_temperature_interactive(data_sw_selected),
    "temperature_surface-water_time-series_selected"
  )
  
  save_to_html(
    plot_temperature_interactive(data_gw), 
    "temperature_groundwater_time-series_full"
  )
  
  save_to_html(
    plot_temperature_interactive(data_gw_selected),
    "temperature_groundwater_time-series_selected"
  )
  
  save_to_html(
    plot_prediction_interactive(predictions), 
    "temperature_prediction"
  )
  
  # save_to_html(
  #   plot_residuals_interactive(prediction_df, binwidth = 0.5),
  #   "temperature_prediction_residuals"
  # )
})  
```
