---
title: "Tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Installation

```{r setup, eval = FALSE}
### Optionally: specify GitHub Personal Access Token (GITHUB_PAT)
### See here why this might be important for you:
### https://kwb-r.github.io/kwb.pkgbuild/articles/install.html#set-your-github_pat

# Sys.setenv(GITHUB_PAT = "mysecret_access_token")

# Install package "remotes" from CRAN
if (! require("remotes")) {
  install.packages("remotes", repos = "https://cloud.r-project.org")
}

# Install KWB package 'kwb.heatsine' from GitHub
remotes::install_github("KWB-R/kwb.heatsine")

```

# Data preparation 

The sinus fit optimisation works for for daily time series of a surface water and 
a groundwater monitoring point, which follow the following requirements: 

## File naming

The time series must be provided in a `.csv` (comma separated value) file, which 
follows the naming convention:

temperature_`<type>`_`<monitoring_id>`.csv

where `<type>` needs to be exactly `groundwater` or `surface-water`. In case any 
other value for `<type>` is chosen the program will not work (e.g `SURFACE WATER`). 
The value of the `<monitoring_id>` will be used for labeling the monitoring points
in the plots and can be value but must not contain `_` (underscore) or `" "` (space).

# Data structure  

The data structure of each `.csv` file is very simple as it contains only the 
two columns `date` (formatted: `YYYY-MM-DD`, e.g. `2020-06-23`) and `value` (temperature 
data in degree Celsius) as shown exemplary below:

```{r eval = FALSE}
date,value
2020-06-23,12.3
2020-06-24,12.3

```

Please make sure that both columns `date` and `value` are all lowercase and correctly 
spelled as the program will not work if this is not the case.

# Data preparation

## Import

```{r}

path_csv_sw <- kwb.heatsine::extdata_file("temperature_surface-water_TEGsee-mikrosieb.csv")
path_csv_gw <- kwb.heatsine::extdata_file("temperature_groundwater_TEG343.csv")


data_sw <- kwb.heatsine::load_temperature_from_csv(path_csv_sw)
data_gw <- kwb.heatsine::load_temperature_from_csv(path_csv_gw)

```

## Plot

Plot time series for whole time period (moveover on data points to select)
  
### Surface water

```{r}
kwb.heatsine::plot_temperature_interactive(data_sw)
```

### Groundwater

```{r}
kwb.heatsine::plot_temperature_interactive(data_gw)
```


# Select

Reduce time period to user input (`date_end` is optional, if not given 
it is set to 'date_start' + 365.25 days)

```{r}

data_sw_selected <- kwb.heatsine::select_timeperiod(data_sw, 
                                                    date_start = "2015-10-10",
                                                    date_end = "2016-10-14")


data_gw_selected <- kwb.heatsine::select_timeperiod(data_gw,
                                        date_start = "2015-12-28",
                                        date_end = "2016-12-26")
  
```

Plot selected datasets:

## Surface Water (selected)

```{r}
kwb.heatsine::plot_temperature_interactive(df = data_sw_selected)
```

## Grundwater Water (selected)

```{r}
kwb.heatsine::plot_temperature_interactive(df = data_gw_selected)
```


# Sinus optimisation 


## Run

```{r}

limits <- c(100,500) # minimum/maximum period length
tolerance <- 0.001 # the desired accuracy ()
debug <- TRUE

sinusfit_sw <- kwb.heatsine::optimise_sinus_variablePeriod(temp_df = data_sw_selected,
                                               opt_limits = limits,
                                               opt_tolerance = tolerance,
                                               opt_debug = debug)

sinusfit_gw <- kwb.heatsine::optimise_sinus_variablePeriod(temp_df = data_gw_selected,
                                               opt_limits = limits,
                                               opt_tolerance = tolerance,
                                               opt_debug = debug)

# Generate data frame with simulated and observed values
predictions <- kwb.heatsine::get_predictions(sinusfit_sw, sinusfit_gw, retardation_factor = 1.8)
```

## Analyse

### Interactive Plot

```{r}
kwb.heatsine::plot_prediction_interactive(predictions)
```


### Table: Traveltimes
```{r}
knitr::kable(predictions$traveltimes)
```

### Table: Goodness of Fit 
```{r}
knitr::kable(predictions$gof)
```

### Table: Optimisation Parameters

```{r}
knitr::kable(predictions$paras)
```

### Table: Data

```{r}
# only first entries
knitr::kable(head(predictions$data))
```


## Export 

Export results to `csv`:

```{r}
kwb.utils::createDirectory("csv")
readr::write_csv(predictions$data,
                 path = "csv/sinus-fit_predictions.csv")
readr::write_csv(predictions$paras,
                 path = "csv/sinus-fit_parameters.csv")
readr::write_csv(predictions$gof,
                 path = "csv/sinus-fit_goodness-of-fit.csv")
readr::write_csv(predictions$traveltimes,
                 path = "csv/sinus-fit_traveltimes.csv")
readr::write_csv(
  data.frame(residuals = predictions$data$simulated - predictions$data$observed),
  "csv/sinus-fit_residuals.csv"
)


## Export plots:
kwb.utils::createDirectory("plots")
withr::with_dir(new = "plots", code = {
  plot_temperature_interactive(data_sw) %>%
    htmlwidgets::saveWidget("temperature_surface-water_time-series_full.html")
  
  plot_temperature_interactive(df = data_sw_selected) %>%
    htmlwidgets::saveWidget("temperature_surface-water_time-series_selected.html")
  
  plot_temperature_interactive(data_gw) %>%
    htmlwidgets::saveWidget("temperature_groundwater_time-series_full.html")
  plot_temperature_interactive(df = data_gw_selected) %>%
    htmlwidgets::saveWidget("temperature_groundwater_time-series_selected.html")
  
  plot_prediction_interactive(predictions) %>%
    htmlwidgets::saveWidget("temperature_prediction.html")
  
  # plot_residuals_interactive(prediction_df, binwidth = 0.5) %>%
  #   htmlwidgets::saveWidget("temperature_prediction_residuals.html"
})  
```
